---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{R}
getdata <- function () {

  myZip = "./activity.zip"
  csvpath="./temp/activity.csv"
  #unzip the data package into a temp directory
  unzip(myZip, exdir="./temp")

  mydata <- read.csv(csvpath)
  mydata$date <- as.Date(mydata$date)
  
  #head(mydata)
  
  return(mydata)
}

  mydata <- getdata()
  
stepshistogram <- function(mydata){
  steps.sum <- tapply(mydata$steps, mydata$date, FUN=sum)
  steps.sum.dataframe <- as.data.frame(steps.sum)
  hist(steps.sum.dataframe$steps, xlab="Day", ylab="Total steps")
  return(steps.sum)
  }
  
stepshistogram(mydata)
  
```

## What is mean total number of steps taken per day?

```{r}
get.means.data <- function(mydata){
  means.data <- tapply(mydata$steps, mydata$date, FUN=mean, na.rm=TRUE)
  return(means.data)
}
get.medians.data <- function(mydata) {
  medians.data <- tapply(mydata$steps, mydata$date, median, na.rm=TRUE)
  #sapply(mydata, function(i) tapply(df[[i]], df$group, sd, na.rm=TRUE))
  return(medians.data)
}

means.data <- get.means.data(mydata)
medians.data <- get.medians.data(mydata)
means.medians.matrix <- cbind(means.data, medians.data)

print(means.medians.matrix)
```

## What is the average daily activity pattern?
What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
plot.daily.activity <- function(mydata) {
  daily.activity.data <- tapply(mydata$steps, mydata$interval, FUN=mean, na.rm=TRUE)
  class(daily.activity.data)
  
  plot(daily.activity.data, type="l", main='Daily Activity', xlab = "Interval", ylab="Average number of steps")
  return(daily.activity.data)
}


plot.daily.activity(mydata)
```

## Imputing missing values





## Are there differences in activity patterns between weekdays and weekends?

This code creates a panel plot containing a time series plot of the 5-minute interval during the day and the average number of steps taken, averaged across all weekday days or weekend days. 

```{r}
weekend.activity <- function(mydata) {
  # Create a factor with the day of the week
  day <- weekdays(mydata$date) 
  alldays.matrix <- cbind(mydata, day)
  # Replace the day of the week with a weekend or weekday factor
  alldays.matrix$day <- as.character(alldays.matrix$day)
  alldays.matrix$day[alldays.matrix$day == "Monday" |  alldays.matrix$day == "Tuesday" | 
                        alldays.matrix$day == "Wednesday" | alldays.matrix$day == "Thursday"| 
                        alldays.matrix$day == "Friday"] <- "weekday"

  alldays.matrix$day[alldays.matrix$day == "Sunday" | alldays.matrix$day == "Saturday" ] <- "weekend"
  
  weekdays.matrix <- subset(alldays.matrix, day == "weekend")
  weekend.matrix <- subset(alldays.matrix, day == "weekday")
  
  #plot
  weekend.activity.data <- tapply(weekend.matrix$steps, weekend.matrix$interval, FUN=mean, na.rm=TRUE)
  weekday.activity.data <- tapply(weekdays.matrix$steps, weekdays.matrix$interval, FUN=mean, na.rm=TRUE)
  
  par(mfrow=c(2,1))
  plot(weekend.activity.data, type="l", xlab="Interval", ylab="average steps", main="Weekends")
  plot(weekend.activity.data, type="l", xlab="Interval", ylab="average steps", main="Weekdays")
} 

weekend.activity(mydata)

```
