---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{R}
getdata <- function () {

  myZip = "./activity.zip"
  csvpath="./temp/activity.csv"
  #unzip the data package into a temp directory
  unzip(myZip, exdir="./temp")

  mydata <- read.csv(csvpath)
  mydata$date <- as.Date(mydata$date)
  
  #head(mydata)
  
  return(mydata)
}

  mydata <- getdata()
  
stepshistogram <- function(mydata){
#   steps.sum <- tapply(mydata$steps, mydata$date, FUN=sum)
#   steps.sum.dataframe <- as.data.frame(steps.sum)
#   hist(steps.sum.dataframe$steps, xlab="Day", ylab="Total steps")
  
   #make a histogram
  library(ggplot2)
  p <- ggplot(data=mydata, aes(mydata$steps)) 
  p <- p + geom_histogram()
  print(p)
  
  #return(steps.sum)
  }
  
stepshistogram(mydata)
  
```

## What is mean total number of steps taken per day?

```{r}
get.means.data <- function(mydata){
  means.data <- tapply(mydata$steps, mydata$date, FUN=mean, na.rm=TRUE)
  return(means.data)
}
get.medians.data <- function(mydata) {
  medians.data <- tapply(mydata$steps, mydata$date, median, na.rm=TRUE)
  #sapply(mydata, function(i) tapply(df[[i]], df$group, sd, na.rm=TRUE))
  return(medians.data)
}

means.data <- get.means.data(mydata)
medians.data <- get.medians.data(mydata)
means.medians.matrix <- cbind(means.data, medians.data)

print(means.medians.matrix)

```
Provide the summary of this data for comparison against the later question with imputed data.
```{r}
summary(means.medians.matrix)
```

## What is the average daily activity pattern?
What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
plot.daily.activity <- function(mydata) {
  daily.activity.data <- tapply(mydata$steps, mydata$interval, FUN=mean, na.rm=TRUE)
  class(daily.activity.data)
  
  plot(daily.activity.data, type="l", main='Daily Activity', xlab = "Interval", ylab="Average number of steps")
}


plot.daily.activity(mydata)
```

## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing data filled in.
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
impute.activity.data <- function(mydata) {
  # Get a data frame of the average number of steps during each interval, regardless of day

  imputed.activity.data <- mydata
  # debug step
  cat("The unimputed matrix")
  print(head(imputed.activity.data))
  # identify the NA values in the original data and replace with the mean of all steps
  imputed.activity.data$steps[which(is.na(imputed.activity.data$steps))] <- mean(imputed.activity.data$steps, na.rm = TRUE)
  # debug step
  cat("The  matrix with imputed data")
  print(head(imputed.activity.data))

    #make a histogram
  library(ggplot2)
  p <- ggplot(data=imputed.activity.data, aes(imputed.activity.data$steps)) 
  p <- p + geom_histogram()
  print(p)
  
#   means.data <- get.means.data(imputed.activity.data)
#   medians.data <- get.medians.data(imputed.activity.data)
#   imputed.means.medians.matrix <- cbind(means.data, medians.data)
  
  #print(imputed.means.medians.matrix)
  return(imputed.activity.data)
}

imputed.activity.data <- impute.activity.data(mydata)
means.data <- get.means.data(imputed.activity.data)
medians.data <- get.medians.data(imputed.activity.data)
imputed.means.medians.matrix <- cbind(means.data, medians.data)
print(imputed.means.medians.matrix )

```

The summary of this data for comparison against data with missing values.
```{r}
summary(imputed.means.medians.matrix)
```


## Are there differences in activity patterns between weekdays and weekends?

This code creates a panel plot containing a time series plot of the 5-minute interval during the day and the average number of steps taken, averaged across all weekday days or weekend days. 

```{r}
weekend.activity <- function(mydata) {
  # Create a factor with the day of the week
  day <- weekdays(mydata$date) 
  alldays.matrix <- cbind(mydata, day)
  # Replace the day of the week with a weekend or weekday factor
  alldays.matrix$day <- as.character(alldays.matrix$day)
  alldays.matrix$day[alldays.matrix$day == "Monday" |  alldays.matrix$day == "Tuesday" | 
                        alldays.matrix$day == "Wednesday" | alldays.matrix$day == "Thursday"| 
                        alldays.matrix$day == "Friday"] <- "weekday"

  alldays.matrix$day[alldays.matrix$day == "Sunday" | alldays.matrix$day == "Saturday" ] <- "weekend"
  
  weekdays.matrix <- subset(alldays.matrix, day == "weekend")
  weekend.matrix <- subset(alldays.matrix, day == "weekday")
  
  #plot
  weekend.activity.data <- tapply(weekend.matrix$steps, weekend.matrix$interval, FUN=mean, na.rm=TRUE)
  weekday.activity.data <- tapply(weekdays.matrix$steps, weekdays.matrix$interval, FUN=mean, na.rm=TRUE)
  
  par(mfrow=c(2,1))
  plot(weekend.activity.data, type="l", xlab="Interval", ylab="average steps", main="Weekends")
  plot(weekend.activity.data, type="l", xlab="Interval", ylab="average steps", main="Weekdays")
} 

weekend.activity(mydata)

```
